% !TEX root = main.tex

\section{Translation Tool}

The iUML-B tooling is based on the Eclipse Modelling Framework (EMF). It is therefore beneficial to load the SCXML model into EMF so that our existing model to model 
transformation technology can be used to define the new SCXML to iUML-B translation. An EMF meta-model for SCXML is available from the Sirius \cite{??}
project. It supports SCXML functionality as well as providing generic model loading facilities for new namespace extensions such as those we introduce in \ref{sect:extension}.

Hierarchical nested state charts are translated to similar  corresponding state-machine structures in iUML-B. 
There are two alternative styles of Event-B representation for iUML-B state-machines.  
Currently the state-variables style is adopted because it is simpler to translate from the Scxml model. 
However, the alternative state-enumeration style has benefits in user readability and may be supported in future.
This would  require conventions regarding the name of the state-machine  to be adopted and used by the modeller in order to construct guards that refer to the current value of the state-machine.
SCXML features, such as initial states, entry/exit actions, and transition actions have corresponding similar features in iUML-B and their translation is relatively straightforward. Since SCXML `final' states, unlike iUML-B final states, are more akin to real states, their translation is less straightforward. An iUML-B state, final state and transition from the former to the latter are added to the corresponding iUML-B state-machine. The transition elaborates all Event-B events that are elaborated by transitions that exit the parent iUML-B state. 

\subsection{Refinement Levels}
An \emph{iumlb:refinement} attribute is used to indicate the first refinement level at which an element should be introduced in the generated iUML-B/Event-B model. In general, elements with no refinement attribute adopt that of their parent.  However, for \emph{iscxml:State elements}, the refinement level refers to any state 
machines generated from the children nested in that state irrespective of whether those children specify a different refinement level. This is because generated iUML-B states cannot be added to an existing state-machine in later refinements.
For \emph{iumlb:invariants} the generated invariant is only  generated at the specified refinement level, not in  subsequent refinements. This is because Event-B invariants are visible through all subsequent refinements.

Note that our approach to refinement in SCXML largely restricts us to superposition refinement where entirely new details are introduced at each refinement level.  It may be possible to support ranges in the refinement attribute enabling a model element to be replaced by some other element in a true data refinement. We plan to investigate this in future work, although, clearly, several coexisting alternative representations of the same concept may be problematic for the SCXML semantics.

\subsection{Constructing event elaborated by transitions}
The Event-B events that are elaborated by an iUML-B  transition are named as follows. 

\begin{enumerate}
\item If the transition has \emph{iumlb:label} attributes, events are generated and named according to the label attributes.
\item If the transition's source is an initial state at the outer state  chart level the transition elaborates the special Event-B INITIALISATION event. 
\item If the transition's source is an initial state of a nested state  chart the names of all the events that are associated with  incoming transitions to the parent state are used.
\item if none of the above provide any labels, a default  `source\_target' format is used.
\end{enumerate}

Trigger events are deliberately not used for  transition events because we want to keep them as a 
separate concept from transition firing in line with SCXML semantics.

\subsection{Data elements}
Data elements, collated in Datamodel elements, model ancillary variables in the way usual SCXML style. Data elements are translated to Event-B variables of type given in an \emph{iumlb:type} attribute translated into an Event-B subset invariant.  The id attribute of the data element is interpreted as the name of the variable and the value is used as the right hand side of an assignment action to initialise the variable.  Some syntax conversion is performed to convert the predicate from SCXML  format into the Event-B mathematical language. The variable is introduced at the same refinement level as the parent element that contains it.


%------------------------------------------------------------------------------

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "HDMachine"
%%% End: 
