% !TEX root = ../main.tex


\section{Introduction}
\label{sec:intro}

\KarlaInlineComment{New context for journal article:}
\KarlaInlineComment{ We simulate the translation with the scenario checker. It automatically runs to completion (events are annotated as internal and prioritised)
 This shows the main intended R2C behaviour for a refinement level.
 Other behaviours are possible depending on future refinement - e.g. a completion can occur without taking enabled transitions if in future we might strengthen a guard.
 	these can be explored by single event stepping without the scenario checker big step.}

\KarlaInlineComment{ Implementing a queue for the triggers }

\KarlaInlineComment{ Proving liveness properties instead of model checking it }
\ColinInlineComment{Do we still discuss the difficulty of refinement and R2C? If not should we add it back in?}

Reactive \SCs are open systems capable of receiving potentially non-deterministic input. 
%This work exposes a shallow embedding of open \SCs~\cite{MoSn16,MoSnHo18,MoSnHo-ABZ2020} semantics in \EventB.
\SCs provide a graphical language, generalized from state machines, that is popular with engineers.
Variants appear in Matlab Simulink/Stateflow~\cite{MATLAB:2019} and the Ansys tools.  
Particularly attractive is providing accessibility to abstraction/refinement via Rodin/\EventB which has an intuitive metaphor in the \SC semantics~\cite{MoSnHo18,MoSnHo-ABZ2020}.  
%The commercial tools have similar ideas expressed as encapsulation and composition but not entailing any formal guarantees. 
The hope is that engineers can better understand the origin of proof obligations in refinements and achieve formal guarantees earlier in their designs where it is most tractable.
%Other researchers have developed a number of refinement semantics for different purposes~\cite{Syriani_2019,Harel,Maraninchi91theargos}.  
Our approach is focused on a mapping to \EventB where safety preservation is key.  
%\EventB provides not only a definition of refinement but a rubric for finding valid refinements and this is carried over into the \SCs work presented here. 
In our version of \SC semantics, refinement means a subset of traces from an abstraction. 
This has the beneficial effect of preserving safety properties from abstraction to refinement and permits proofs to be discharged at the highest tractable level of abstraction where they are the easiest to discharge.

Many incompatible definitions of refinement have been posed by others~\cite{Syriani_2019,Maraninchi91theargos} and that can lead to confusion.  
Though these separate refinements have different goals, all of which may be attractive to systems designers in different ways,
% preservation of safety proofs is the goal here. 
%While these other forms of refinement cannot be said to conflict with the one presented here, 
they will not always preserve safety properties.  
From the \EventB vernacular it might be better to relabel these other approaches not as methods of model ``refinement'', but rather methods of model ``elaboration''.  
Preservation of safety across refinement requires only a few restrictions to the original~\cite{Harel} \SCs (e.g. transitions cannot cross containment boundaries arbitrarily), but still allows for both parallel and hierarchical composition. 
%With these restrictions composition becomes a refinement, but not all refinements are compositions.  
%Such a unification of composition and refinement can lead, not only to code reuse, but reuse of proofs.
\input{sections/relatedWork}

\ColinInlineComment{Check outline below is correct before submission}
%Section~\ref{sec:relatedWork} discusses related work on refinement and what are the different types.
Section~\ref{sec:background} provides background material. 
Section~\ref{sec:run-completion} discusses the \SC concept of `run to completion' and how it can be specified in \EventB. 
Section~\ref{sec:descr-sample-appl} introduces our example case study; a drone. 
Section~\ref{sec:translation} gives an outline of our translation from \SCXML to \EventB via \UMLB.
Section~\ref{sec:validation} describes the use of \UMLB animation and Scenario Checking tools to validate translated \SCXML models.
Section~\ref{sec:verificationSafety} illustrates our approach to verifying safety invariant properties. 
Section~\ref{sec:verificationResponses} illustrates our approach to verifying control responses, and 
Section~\ref{sec:conc} concludes.

%% {\color{red}
%% Ideas on content
%% We prove properties in early refinements - because easier to see - (so far fully automatic - try proving battery inv in later refinement)
%% 	although we put everything in the same model we envisage gradual construction	
%% 	we could also build a Sirius tool that can show views at each refinement.

%% proof forces us to finalise transitions..
%% 	.. i.e. to remove non-determinism that allows for some future refinement.
%% 		(in usual event-b it is not built in to the 'engine' and is only checked when the next refinement is provided..
%% 		.. we only have this finalisation because we needed to model the run to completion queueing mechanisms, hence new transitions modify old variables (queues))
%% 	we have to limit what can be done in later refinements as some things could violate what is proved.
%% 	e.g. a response transition must come next (to make the invariant before completion) so we cannot strengthen its guards any more.
%% 	e.g. similarly a trigger raised as a response must be used in the way that preserves the invariant, so future transitions may not consume it.

%% Also section on how we arrived at the new translation.
%% 	adding the invariant about only one dequeued trigger at a time - otherwise proof did not discharge about the future triggered transitions
%% 		(because consuming another dequeued trigger would discard the one we are interested in)
%% 	the drone model made us rethink the basis.. we had to dequeue triggers.. (why?)
%% }

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: